---
title: "Projet 8 - Master Regulators for Metastatic behavior in Osteosarcoma"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chabrier Lisa, Cortes Garcia Estaban, Dorval Loïc

# I- Introduction
TODO : Lecture d'article, intro du sujet.

# II- Analysis
Project step by step:
- You should download the gene expression data generated by the study (GSE21257);

- You should install the RTN package and infer the regulatory network based on the down-
loaded gene expression data;

- You should install snow package for parallel processing in RTN;
- You should install RedeR package to better visualize the network you inferred;

- By performing a differential gene expression analysis between the metastatic and the non-
metastatic biopsies, you will obtain a signature for the metastatic behavior of this cancer

in these patients;
- You should run the Master Regulator Analysis to infer putative Master Regulators by using
the inferred network and the obtained signature.
- Get biological insight into these putative master regulators at the Human Protein Atlas
and at Gene Cards.

# Exemples : 
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
